---
title: "Cleaning Biochemical Data"
author: "Camila Vargas"
date: "10/20/2021"
output: html_document
---


This scripts created matlab files into csvs and documents David's Koweek Biochemical data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library (plyr)
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(R.matlab)


## Sets R not to use scientific notations
options(scipen=999) 
```


## Read file
```{r}
biochem_raw <- readMat(here::here("6.raw_data/Koweek et al_Biogeochemical_Data.mat"))
```

## Convert to DF
```{r}
## rbinds all objects on the list and keeps the list name as .id
df <- ldply (biochem_raw, data.frame)

```

There are 5 tubes.
For each tubes we have:

- Time: time in standard MATLAB format (days)
-S: salinity (psu)-data presented in Fig. 3
- DOuM: dissolved oxygen (uM)
- pH_calc: pH (total scale) calculated from DIC and TA
- pCO2: pCO2 (uatm)
- DIC: dissolved inorganic carbon (umol kg^-1)
- sDIC: DIC salinity-normalized to 34.87 (umol kg^-1)
- TA: total alkalinity (umol kg^-1)
- sTA: TA salinity-normalized to 34.87 (umol kg^-1)
- Omega_Ar: aragonite saturation state calculated from DIC and TA

## Oganizing and renaming

```{r}

biochem_long <- df %>% 
  separate(col = ".id",
           into = c("variable", "col2", "date"),
           sep = "\\.") %>% 
  dplyr::mutate(date = ifelse(is.na(date), col2, date)) %>% 
  dplyr::rename(
         tube_1 = X1,
         tube_2 = X2,
         tube_3 = X3,
         tube_4 = X4,
         tube_5 = X5) %>% 
  select(-col2) %>% 
  pivot_longer(cols = 3:7,
               names_to = "sample_tube")


time_cols <- biochem_long %>% 
  filter(variable == "Time") %>% 
   mutate(time = as.POSIXct((value-719529)*86400, origin = "1970-01-01", tz = "HAST")) %>% 
  select(-value)


biochem_clean <- biochem_long %>% 
  filter(variable != "Time") %>% 
  bind_rows(time_cols) %>% 
  
  pivot_wider(id_cols = c(time, sample_tube, date),
              values_from = value,
              names_from = variable)
  
  

```

## testing
```{r}

df_time <- ldply (biochem_raw[22], data.frame) %>% 
  mutate(X1 = as.POSIXct((X1-719529)*86400, origin = "1970-01-01", tz = "HAST"))

```

