---
title: "Cleaning PCA data"
author: "Camila Vargas"
date: "10/21/2021"
output: html_document
---

This scripts created matlab files into csvs and documents David's Koweek PCA data.

3. Metabolic rate data from Lagrangian and Eulerian methods along with associated environmental and ecological conditions
File Name: Koweek et al_PCA_data.mat

Contains the following variables:

**NCC:** Net community calcification (mmol m^-2 hr^-1)
**NCC_SD:** 1 standard deviation error estimate on NCC (mmol m^-2 hr^-1)
**NCP:** Net community production (mmol m^-2 hr^-1)
**NCP_SD:** 1 standard deviation error estimate on NCP (mmol m^-2 hr^-1)
**NCC_to_NCP:** Ratio of NCC to NCP
**Fractional_Day:** Fractional time of day (days; e.g., 0600 hrs=0.25 days, 1200 hrs=0.5 days, 1800 hrs=0.75 days)
**Water_Temperature:** Average water temperature during each metabolic rate measurement (deg C)
**Air_Temperature:** Air temperature from meteorological station on Palmyra Atoll (deg C)
**PAR:** photosynthetically active radiation from meteorological station on Palmyra Atoll (μmol m^-2 s^-1)
**Precipitation:** precipitation from meteorological station on Palmyra Atoll (mm hr^-1)
**Hard_Coral:** Average percentage of hard coral during either Lagrangian Drift or over flow path from fore reef to SIB site for Eulerian measurements (%)
**Macroalgae:** Average percentage of macroalgae during either Lagrangian Drift or over flow path from fore reef to SIB site for Eulerian measurements (%)
**CCA:** Average percentage of crustose coralline algae during either Lagrangian Drift or over flow path from fore reef to SIB site for Eulerian measurements (%)
**Halimeda:** Average percentage of Halimeda spp. during either Lagrangian Drift or over flow path from fore reef to SIB site for Eulerian measurements (%)
**Turf_Algae:** Average percentage of turf algae during either Lagrangian Drift or over flow path from fore reef to SIB site for Eulerian measurements (%)
**Calc_to_NonCalc:** Ratio of calcifiers to non-calcifiers during either Lagrangian Drift or over flow path from fore reef to SIB site for Eulerian measurements
**Hard_Coral_to_Macroalgae:** Ratio of hard coral to macroalgae during either Lagrangian Drift or over flow path from fore reef to SIB site for Eulerian measurements


All variables are concatenated in the variable "All_data" according to the sequence listed in the variable "Column_labels_All_data".

Note: The variable All_data (after removing "NCC_SD" and "NCP_SD") was Z-standardized to compute the PCA (Fig. 6). The covariance matrix on the Z-standardized data set was computed using 'pairwise' methods (i.e., Cov(i,j) was calculated using all rows in columns i and j for which data exists).


## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library (plyr)
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(R.matlab)

## Sets R not to use scientific notations
options(scipen=999) 
```

## Read data
```{r}

pca_raw <- readMat(here::here("6.raw_data/Koweek et al_PCA_data.mat"))

```

## Convert to DF
```{r}
## rbinds all objects on the list and keeps the list name as .id
pca_df <- ldply (pca_raw, data.frame)

```

## Row names

From David's README:
"Row_labels" lists the corresponding Lagrangian Drift (LD) number or time of Eulerian rate measurement at the SIB site. 
```{r}

ld_number <- filter(pca_df, .id == "Row.labels") %>% 
  select(37:100) %>% 
  pivot_longer(cols = everything(),
               names_to = "list_name",
               values_to = "ld_SIB_row") %>% 
  dplyr::mutate(padl_id = 1:dplyr::n()) %>% 
  select(-list_name)

## 64 rows. 16 LD and 48 Eulerian rate measurment
```

## Colnames

```{r}

names_col <- filter(pca_df, .id == "Column.labels.All.data") %>% 
  select(20:36) %>% 
  pivot_longer(cols = everything(),
               names_to = "list_names",
               values_to = "colnames")

col_names_vec <- pull(names_col, "colnames")
```

## All data
Pulling out all concatenated data from master DF

```{r}

all_data <- pca_df %>% 
  filter(.id == "All.data") %>% 
  select(2:18) %>% 
  rename_at(vars(everything()), ~col_names_vec ) %>% 
  dplyr::mutate(padl_id = 1:dplyr::n())

```

## Complete data set with names

```{r}
pca_clean <- all_data %>% 
  left_join(ld_number, by = "padl_id") %>% 
  select(LD_Eu = ld_SIB_row, everything(1:17))
```


**NEED TO CHECK WITH SUPPLEMENTSRY MATERIAL**
Make sure that the order in all data matches the order in the row names and col names.


## Saving data
```{r}

write_csv(pca_clean, here::here("7.clean_data/2012_Koweek_et_al_pca_data.csv"))

```

