---
title: "Cleaning Biochemical Data"
author: "Camila Vargas"
date: "10/20/2021"
output: html_document
---


This scripts created matlab files into csvs and documents David's Koweek Biochemical data.

Biogeochemical data from Eulerian measurements at the SIB site

File Name: Koweek et al_Biogeochemical_Data.mat

Column numbers correspond with tube numbers at the SIB site (i.e., column 1 in each variable corresponds with data collected from tube 1, column 2 corresponds with tube 2, etc.)
Each variable matrix ends in suffixes "_20Sep", "2122Sep", or "25Sep" corresponding to data collected on 20 September 2012, 21-22 September 2012, and 25 September 2012, respectively.

Contains the following variables:

**Time:** time in standard MATLAB format (days)
**S:** salinity (psu)-data presented in Fig. 3
**DOuM:** dissolved oxygen (uM)
**pH_calc:** pH (total scale) calculated from DIC and TA
**pCO2:** pCO2 (uatm)
**DIC:** dissolved inorganic carbon (umol kg^-1)
**sDIC:** DIC salinity-normalized to 34.87 (umol kg^-1)
**TA:** total alkalinity (umol kg^-1)
**sTA:** TA salinity-normalized to 34.87 (umol kg^-1)
**Omega_Ar:** aragonite saturation state calculated from DIC and TA


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library (plyr)
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(R.matlab)
library(freeR)


## Sets R not to use scientific notations
options(scipen=999) 

```


## Read file
```{r}
biochem_raw <- readMat(here::here("6.raw_data/Koweek et al_Biogeochemical_Data.mat"))
```

## Convert to DF
```{r}
## rbinds all objects on the list and keeps the list name as .id
df <- ldply (biochem_raw, data.frame)

```
Each X column is a tube
.id has date and variable name

## Oganizing and renaming

```{r}
biochem_long <- df %>% 
  separate(col = ".id",
           into = c("variable", "col2", "date"),
           sep = "\\.") %>% 
  dplyr::mutate(date = ifelse(is.na(date), col2, date)) %>% 
  dplyr::rename(
         tube1 = X1,
         tube2 = X2,
         tube3 = X3,
         tube4 = X4,
         tube5 = X5) %>% 
  arrange(variable, date) %>% 
  dplyr::mutate(position = rep(1:120, times = 10)) %>% 
  select(-col2) %>% 
  pivot_longer(cols = 3:7, ##all tube columns
               names_to = "tube")


time_data <-  biochem_long %>% 
  filter(variable == "Time") %>% 
  mutate(Time = as.POSIXct((value-719529)*86400, origin = "1970-01-01", tz = "HAST")) %>%
  select(-value, -variable)

biochem_wide <- biochem_long %>% 
  filter(variable != "Time") %>% 
  pivot_wider(id_cols = c(date, position, tube),
              names_from = "variable",
              values_from = "value") %>% 
  left_join(time_data, by = c("date", "position", "tube")) %>% 
  select(Time, date, everything(), -position)

```

Something is not matching! Biochem_all should be length 1200?? Is there data missing from date Sept2122?

Re check math and what can be happening.
Also, when plotting the data Salinity ans DIC for days 21 and 25 do not plot because they do not have a time associated. 
WHAT IS HAPPENING??

The way we were assigning the position was not right that is why we were getting a longer data set than expected.
Problem fixed!

The wide data frame has a length of 600. This means 120 samples (17 Sept 20, 79 Sept 21-22 and 24 Sept 25) times 5 tubes.

## Checknig
```{r}
test <- biochem_wide %>% 
  group_by(date) %>% 
  tally()

View(test)

test_2 <- biochem_wide %>% 
  group_by(tube) %>% 
  tally()

View(test_2)

test_3 <- biochem_wide %>% 
  group_by(date,tube) %>% 
  tally()

View(test_3)

```



## Re-creating papers fig4 to check
```{r}


plot_data <- biochem_long %>%
  filter(variable != "Time") %>% 
  left_join(time_data, by = c("position", "tube", "date"))
  
ggplot(plot_data, 
       aes(x = Time, y = value, color = tube)) +
           geom_point() +
  facet_grid(variable ~ ., scales = "free") +
           #facet_wrap(~variable, scales = "free") +
           theme_bw()
  

```

Looks good!!!


## Save data
```{r}
write_csv(biochem_wide, here::here("7.clean_data/2012_Koweek_biogeochemical_data.csv"))
```


