---
title: "Cleaning Physical Data"
author: "Camila Vargas"
date: "10/22/2021"
output: html_document
---
Information about this data set from David Koweek's README:

1. Physical oceanographic data from Eulerian measurements at the SIB site
File Name:  Koweek et al_Physical_Data.mat

Column numbers correspond with tube numbers at the SIB site (i.e., column 1 in each variable corresponds with data collected from tube 1, column 2 corresponds with tube 2, etc.)

Contains the following variables:

**Eta_avg:** time-averaged free surface at SIB site (m)
**PAR:** photosynthetically active radiation from meteorological station on Palmyra Atoll (μmol m^-2 s^-1)
**Temp_bot:** bottom temperature (deg C)
**Temp_mid:** temperature at mid water column (only at tubes 2 and 4) (deg C)
**Temp_top:** near-surface water temperature (only at tubes 2 and 4) (deg C)
**U_heading:** compass heading direction of water velocity (deg)
**U_mag:** magnitude of water velocity (m s^-1)
**airT:** air temperature from meteorological station on Palmyra Atoll (deg C)
**density:** average water density across SIB site (kg m^-3)
**depth:** depth (m)
**mtime:** time vector in standard MATLAB format (days)
**precip:** precipitation from meteorological station on Palmyra Atoll (mm hr^-1)
**site_lat:** latitude of tubes (deg)
**site_lon:** longitude of tubes (deg)
**site_names:** names of all five tubes
**tr_fr_sib:** travel time from forereef to SIB tubes (min)
**tr_fr_sd:** travel time from foreeef to start of drift 16 (min)
**wind_dir:** wind direction at 2m from NASA MERRA global model (deg)
**wind_mag:** wind speed at 2m from NASA MERRA global model (m s^-1)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library (plyr)
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(R.matlab)

## Sets R not to use scientific notations
options(scipen=999) 


```


## Read data
```{r}

physical_raw <- readMat(here::here("6.raw_data/Koweek et al_Physical_Data.mat"))

```

## Convert to DF
```{r}
## rbinds all objects on the list and keeps the list name as .id

# tube_num <- c("id", paste0("tube_", 1:5))

physical_df <- ldply (physical_raw, data.frame) %>% 
  select("id" = ".id", tube1 = X1, tube2 = X2, tube3 = X3, tube4 = X4, tube5 = X5, one_var = "X..i..")
  
```


## Tube data
Filtering data that was collected from each tube. This means all variables that have data in columns 1 to 5.

```{r}

tube_var <- c("depth", "U.mag", "U.heading",  "Temp.bot","U.heading", "Temp.bot", "Temp.mid", "Temp.top", "tr.fr.sib")  ##9 variables


tube_data <- physical_df %>% 
  filter(id %in% tube_var) %>% 
  mutate(position = rep(1:692, times = 7)) %>% 
  select(-one_var) %>% 
  pivot_longer(cols = 2:6,
               names_to = "tube",
               values_to = "measurement") %>% 
  pivot_wider(id_cols = c(position, tube),
            names_from = id,
            values_from = "measurement")
  
 
   # group_by(id) %>% 
   # tally()
## Each variable as 692 observations
```



## Context data
Filtering variables that are not directly measure by tube samples, this means they provide information about the setting and is the same across all tubes.

```{r}
one_variable <- c("Eta.avg", "airT", "precip", "PAR", "wind.mag", "wind.dir", "density", "tr.fr.sd") ##8 variables

one_var_data <- physical_df %>% 
  dplyr::filter(id %in% one_variable) %>% 
  mutate(position = rep(1:692, times = 8)) %>% 
  select(id, one_var, position) %>% 
  pivot_wider(id_cols = position,
            names_from = id,
            values_from = one_var)
  

```

## mtime: time vector in standard MATLAB format (days)
Timestamp for each of the 692 samples.
We need to transform this into normal time

```{r}

time_data <- physical_df %>% 
  filter(id == "mtime") %>% 
  mutate(timestamp = as.POSIXct((one_var-719529)*86400, origin = "1970-01-01", tz = "HAST"),
         position = 1:692) %>% 
   select(timestamp, position)

```


## Combining all data
```{r}

one_var_complete <- time_data %>% 
  left_join(one_var_data, by = "position")


all_data <- tube_data %>% 
  left_join(one_var_complete, by = "position") %>% 
  select(timestamp, tube, everything(), -position)

```


## Lat long for each tube
Filtering geographic coodinate for eacha of the sample sites.

```{r}
lat_long <- c("site.lat",  "site.lon")


lat_long_data <- physical_df %>% 
  filter(id %in% lat_long) %>% 
  select(-one_var) %>% 
  pivot_longer(cols = 2:6,
               names_to = "tube",
               values_to = "value") %>% 
  pivot_wider(id_cols = tube,
              names_from = id,
              values_from = value)

```

Physical data has been compile into one dataframe. A second file will be updated with the coords for each tube sampling.

**Next step: Check with Electronic supplementary material that compiling method is accurate.**

## Saving clean data

```{r}
write_csv(all_data, here::here("7.clean_data/2012_Koweek_et_al_physical_data.csv"))

write_csv(lat_long_data, here::here("7.clean_data/2012_Koweek_et_al_eulerian_SIB_coords.csv"))

```



