---
title: "Cleaning Langrangian Drift coordinates"
author: "Camila Vargas"
date: "10/22/2021"
output: html_document
---
Lagrangian Drift tracks
File Name: Koweek_et_al_Lagrangian_Drift_coordinates.mat


Organized into 22-row cell arrays. Position of each cell in the array corresponds with equally-numbered Lagrangian Drift (e.g., Data in 4th cell corresponds with Lagrangian Drift 4). Empty cells in the array correspond to Lagrangian Drifts eliminated during data processing (as discussed in the Methods section of the paper).

Contains the following variables:
LD_Latitudes: Latitudes of all points along each Lagrangian Drift
LD_Longitudes: Longitudes of all points along each Lagrangian Drift

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library (plyr)
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(R.matlab)

## Sets R not to use scientific notations
options(scipen=999) 
```

## Read data
```{r}

lag_coor_raw <- readMat(here::here("6.raw_data/Koweek_et_al_Lagrangian_Drift_coordinates.mat"))

```

## Convert to DF
```{r}
col_n <- paste0("n", 1:11)

lat_df <- lag_coor_raw[1] %>% 
  as_tibble() %>% 
  unnest(cols = c(LD.Latitudes)) %>% 
  separate(LD.Latitudes, col_n, sep = "\\,") %>% 
  mutate(LD_num = paste0("LD ", 1:22),
         n1 = ifelse(n1 == "numeric(0)", NA, n1)) %>% 
  pivot_longer(cols = 1:11,
               names_to = "measurment_num",
               values_to = "lat") %>% 
  mutate(lat = str_replace_all(lat, c("c" = "", "\\(" = "", "\\)" = "")),
         lat = str_trim(lat) %>% as.numeric())


long_df <- lag_coor_raw[2] %>% 
  as_tibble() %>% 
  unnest(cols = c(LD.Longitudes)) %>% 
  separate(LD.Longitudes, col_n, sep = "\\,") %>% 
  mutate(LD_num = paste0("LD ", 1:22),
         n1 = ifelse(n1 == "numeric(0)", NA, n1)) %>% 
  pivot_longer(cols = 1:11,
               names_to = "measurment_num",
               values_to = "long") %>% 
  mutate(long = str_replace_all(long, c("c" = "", "\\(" = "", "\\)" = "")),
         long = str_trim(long) %>% as.numeric())


```


## Match lat and long

```{r}
ld_coords_complete <- lat_df %>% 
  left_join(long_df, by = c("LD_num", "measurment_num")) %>% 
  filter(!is.na(lat))
  
mising_ld <- tibble(
  LD_num = c("LD 1", "LD 2", "LD 8", "LD 11", "LD 14", "LD 17"),
  measurment_num = NA,
  lat = NA,
  long = NA)
  
ld_coords_final <- ld_coords_complete %>% 
  bind_rows(mising_ld) %>% 
  arrange(LD_num)

```

## Safe data
```{r}
write_csv(ld_coords_final, here::here("7.clean_data/2012_Koweek_et_al_LD_coords.csv"))
```


